An Embedding Framework for Consistent Polyhedral Surrogates

### 2021-08-12 poly-embed-journal/poly-embed-journal.tex
lovasz tightness proof

\usepackage{float}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{hyperref}       % hyperlinks  %[implicit=false, bookmarks=false]
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage[normalem]{ulem}
\usepackage{thm-restate}

@@

\begin{lemma}
	\label{lem:lovasz-u}
	Let $\hat\ell^f:\hat\R\times\Y\to\reals_+$, where $\hat\R = \{(A,B) \in 2^N \times 2^N: A\cap B=\emptyset\}$, be given by
	\begin{equation}
	\label{eq:lovasz-embeds}
	\hat\ell^f((A,B),S) = L^f(\chi_A + \ones_B,S) = f(A\triangle S\setminus B) + f(A\triangle S\cup B)~.
	\end{equation}
	Then the Lov\'asz hinge $L^f$ embeds $\hat\ell^f$.
\end{lemma}
\begin{proof}
	As observed above, the set $\U$ is in bijection with $\hat \R$, using the transformation $u = \chi_A + \ones_B$.
	As also observed above, we may write $u = \chi_A \odot \ones_{N\setminus B}$ as well.
	Combining Lemma~\ref{lem:lovasz-trim} with \ref{lem:loss-restrict}, we see that $L^f$ embeds $L^f|_\U$.
	It thus only remains to verify the form~\eqref{eq:lovasz-embeds}.
	We have $L^f(u,S) = F(x)$ where $x = \ones - u\odot\chi_S = \ones - \ones_{N\setminus B} \odot \chi_A \odot \chi_S = \ones + \ones_{N\setminus B} \odot \chi_{A\triangle S} = \ones_{B} + 2\ones_{A\triangle S\setminus B}$.
	%\jessiet{This next line lost me.}
	As $(A\triangle S \setminus B) \cup B = A\triangle S \cup B$, the result follows from~\cite[Prop 3.1(h)]{bach2013learning}.
\end{proof}

@@

\begin{lemma}
  Let $X,Y,Z\subseteq N$ such that $Y\cap Z = \emptyset$.
  Then $X\triangle (Y\cup Z) = X\triangle Y \triangle Z$.
\end{lemma}
\begin{proof}
  $X\triangle (Y\cup Z) = X\cup (Y\cup Z) \setminus (X \cap (Y\cup Z))$

  $X\triangle (Y\cup Z) = (X \setminus (Y\cup Z)) \cup ((Y\cup Z) \setminus X)$
\end{proof}

\begin{theorem}\label{thm:lovasz-embed-tight}
  Let $f$ be strictly submodular and strictly increasing.
  Then $L^f$ tightly embeds $\hat\ell^f$.
\end{theorem}
\begin{proof}
  Let $A,B\subseteq N$ be disjoint.
  We will show that $(A,B)$ is the unique $\hat\ell^f$-optimal report for the distribution $p$ given by
  \begin{equation}
    \label{eq:1}
    p(S) =
    \begin{cases}
      2^{-|B|} & S\setminus A \subseteq B \\
      0 & \text{otherwise}
    \end{cases}~.
  \end{equation}
  The condition $S\setminus A \subseteq B$ is equivalent to the existence of some $T\subseteq B$ such that $S = A \cup T$, and this $T$ must be disjoint from $A$ as $B$ is.
  The expected loss for $(A,B)$ under $p$ is therefore
  \begin{align*}
    \hat\ell^f((A,B);p)
    &= \sum_{T\subseteq B} 2^{-|B|} f(A\triangle (A\cup T)\setminus B) + f(A\triangle (A\cup T)\cup B)
    \\
    &= 2^{-|B|} \sum_{T\subseteq B} f(T\setminus B) + f(T\cup B)
    \\
    &= f(\emptyset) + f(B)~.
  \end{align*}

  Now suppose $A',B'\subseteq N$ are disjoint, with $(A',B') \neq (A,B)$.
  Once again we write the expected loss under $p$.
  We will use the observation that, as $A$ and $T$ are disjoint, $A\cup T = A\triangle T$.
  Defining $\hat A = A' \triangle A$, and using the associativity of $\triangle$, we then have $A' \triangle (A\cup T) = \hat A \triangle T$.
  \begin{align*}
    \hat\ell^f((A',B');p)
    &= 2^{-|B|} \sum_{T\subseteq B} f(A'\triangle (A\cup T)\setminus B') + f(A'\triangle (A\cup T)\cup B')
    \\
    &= 2^{-|B|} \sum_{T\subseteq B} f(\hat A\triangle T\setminus B') + f(\hat A\triangle T\cup B')
    \\
    &= \frac 1 2 \; 2^{-|B|} \sum_{T\subseteq B} \biggl( f(\hat A\triangle T\setminus B') + f(\hat A\triangle T\cup B')
    \\[-8pt]
    & \hspace*{85pt} + f(\hat A\triangle (B\setminus T)\setminus B') + f(\hat A\triangle (B\setminus T)\cup B') \biggr)~.
  \end{align*}

  We first suppose $B \subseteq B'$.
  Then $\hat A\triangle T\setminus B' = \hat A \setminus B'$ and $\hat A\triangle T\cup B' = \hat A\cup B'$ for any $T \subseteq B$.
  We conclude
  \begin{align*}
    \hat\ell^f((A',B');p)
    &= \frac 1 2 \; 2^{-|B|} \sum_{T\subseteq B} \biggl( f(\hat A\setminus B') + f(\hat A\cup B') + f(\hat A\setminus B') + f(\hat A\cup B') \biggr)
    \\
    &= f(\hat A\setminus B') + f(\hat A\cup B')~.
  \end{align*}
  We will show either (i) $\hat A\setminus B' \neq \emptyset$ or (ii) $\hat A\cup B' \supsetneq B$ hold, and then apply the fact that $f$ is strictly increasing.
  If (ii) fails, since $B \subseteq B'$, we must have $B = B'$.
  Then $B = B'$ is disjoint from both $A$ and $A'$, and therefore from $\hat A = A'\triangle A$.
  We therefore have $\hat A\setminus B' \neq \emptyset \iff \hat A \neq \emptyset \iff A \neq A'$, which must hold as $B = B'$ yet $(A,B) \neq (A',B')$ by assumption.
  Thus, at least (i) or (ii) hold.
  As $f$ is strictly increasing, we conclude $\hat\ell^f((A',B');p) > f(\emptyset) + f(B) = \hat\ell^f((A,B);p)$.

  Now suppose $B \not\subseteq B'$.
  We claim that there exists $T\subseteq B$ such that one of the following two cases hold:
  \begin{enumerate}
  \item[(i)]
    $\hat A\triangle T\setminus B'$ and 
    $\hat A\triangle (B\setminus T)\cup B'$
    are incomparable;
  \item[(ii)]
    $\hat A\triangle T\cup B'$
    and
    $\hat A\triangle (B\setminus T)\setminus B'$
    are incomparable. 
  \end{enumerate}
  We will then apply strict submodularity to conclude.

  \bigskip
  Take $T = B\setminus B'$.

  If $B' = \emptyset$

\raf{It seems we need to handle the case where $B\setminus B'$ is a singleton, since then we run into the same problem as when we used only $\emptyset$ and $B$.  If larger, we can take $T$ to fragment $B\setminus B'$ and I think we're good...}

  Let $i\in B\setminus B'$.
  If $i\in A'$, then $i\in\hat A$ (as $i\notin A$).

  \bigskip
  For case (i),

>>>>

\usepackage{float}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{lmodern}
\usepackage{hyperref}       % hyperlinks  %[implicit=false, bookmarks=false]
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage[normalem]{ulem}
\usepackage{thm-restate}
\usepackage{array}
\newcolumntype{R}{>{$}r<{$}} % math-mode version of "l" column type
\newcolumntype{C}{>{$}c<{$}} % math-mode version of "c" column type
\newcommand\numberthis{\addtocounter{equation}{1}\tag{\theequation}}

@@

\raf{Add: explanation for why $B$ cannot be a singleton if we want a tight embedding... or maybe prove this first and then prove the stronger one --- downside with that is needing to define yet another discrete loss.}

\begin{lemma}
	\label{lem:lovasz-u}
	Let $\hat\ell^f:\hat\R\times\Y\to\reals_+$, where $\hat\R = \{(A,B) \in 2^N \times 2^N: A\cap B=\emptyset, |B|\neq 1\}$, be given by
	\begin{equation}
	\label{eq:lovasz-embeds}
	\hat\ell^f((A,B),S) = L^f(\chi_A + \ones_B,S) = f(A\triangle S\setminus B) + f(A\triangle S\cup B)~.
	\end{equation}
	Then the Lov\'asz hinge $L^f$ embeds $\hat\ell^f$.
\end{lemma}
\begin{proof}
	As observed above, the set $\U$ is in bijection with $\hat \R$, using the transformation $u = \chi_A + \ones_B$.
	As also observed above, we may write $u = \chi_A \odot \ones_{N\setminus B}$ as well.
	Combining Lemma~\ref{lem:lovasz-trim} with \ref{lem:loss-restrict}, we see that $L^f$ embeds $L^f|_\U$.
	It thus only remains to verify the form~\eqref{eq:lovasz-embeds}.
	We have $L^f(u,S) = F(x)$ where $x = \ones - u\odot\chi_S = \ones - \ones_{N\setminus B} \odot \chi_A \odot \chi_S = \ones + \ones_{N\setminus B} \odot \chi_{A\triangle S} = \ones_{B} + 2\ones_{A\triangle S\setminus B}$.
	%\jessiet{This next line lost me.}
	As $(A\triangle S \setminus B) \cup B = A\triangle S \cup B$, the result follows from~\cite[Prop 3.1(h)]{bach2013learning}.
\end{proof}

@@

\begin{theorem}\label{thm:lovasz-embed-tight}
  Let $f$ be strictly submodular and strictly increasing.
  Then $L^f$ tightly embeds $\hat\ell^f$.
  \raf{Assuming here $|B|\neq 1$}
\end{theorem}
\begin{proof}
  Let $A,B\subseteq N$ be disjoint.
  We will show that $(A,B)$ is the unique $\hat\ell^f$-optimal report for the distribution $p$ given by
  \begin{equation}
    \label{eq:1}
    p(S) =
    \begin{cases}
      2^{-|B|} & S\setminus A \subseteq B \\
      0 & \text{otherwise}
    \end{cases}~.
  \end{equation}
  The condition $S\setminus A \subseteq B$ is equivalent to the existence of some $T\subseteq B$ such that $S = A \cup T$, and this $T$ must be disjoint from $A$ as $B$ is.
  The expected loss for $(A,B)$ under $p$ is therefore
  \begin{align*}
    \hat\ell^f((A,B);p)
    &= \sum_{T\subseteq B} 2^{-|B|} f(A\triangle (A\cup T)\setminus B) + f(A\triangle (A\cup T)\cup B)
    \\
    &= 2^{-|B|} \sum_{T\subseteq B} f(T\setminus B) + f(T\cup B)
    \\
    &= f(\emptyset) + f(B)~.
  \end{align*}

  Now suppose $A',B'\subseteq N$ are disjoint, with $(A',B') \neq (A,B)$.
  Once again we write the expected loss under $p$.
  We will use the observation that, as $A$ and $T$ are disjoint, $A\cup T = A\triangle T$.
  Defining $\hat A = A' \triangle A$, and using the associativity of $\triangle$, we then have $A' \triangle (A\cup T) = \hat A \triangle T$.
  \begin{align*}
    \hat\ell^f((A',B');p)
    &= 2^{-|B|} \sum_{T\subseteq B} f(A'\triangle (A\cup T)\setminus B') + f(A'\triangle (A\cup T)\cup B')
    \\
    &= 2^{-|B|} \sum_{T\subseteq B} f(\hat A\triangle T\setminus B') + f(\hat A\triangle T\cup B')
      \numberthis\label{eq:tightness-loss-rewrite}
  \end{align*}

  We first suppose $B \subseteq B'$.
  Then $\hat A\triangle T\setminus B' = \hat A \setminus B'$ and $\hat A\triangle T\cup B' = \hat A\cup B'$ for any $T \subseteq B$, giving the simpler form $\hat\ell^f((A',B');p) = f(\hat A\setminus B') + f(\hat A\cup B')$.
  Consider two cases (i) $\hat A\setminus B' \neq \emptyset$, and (ii) $\hat A\cup B' \supsetneq B$.
  We will show either (i) or (ii) hold, and then apply the fact that $f$ is strictly increasing.
  If (ii) fails, since $B \subseteq B'$, we must have $B = B'$.
  Then $B = B'$ is disjoint from both $A$ and $A'$, and therefore from $\hat A = A'\triangle A$.
  We therefore have $\hat A\setminus B' \neq \emptyset \iff \hat A \neq \emptyset \iff A \neq A'$, which must hold as $B = B'$ yet $(A,B) \neq (A',B')$ by assumption.
  Thus, at least (i) or (ii) hold.
  As $f$ is strictly increasing, we conclude $\hat\ell^f((A',B');p) > f(\emptyset) + f(B) = \hat\ell^f((A,B);p)$.

  Now suppose $B \not\subseteq B'$.
  In this case, we will apply strict submodularity.
  Let us rewrite the expected loss,
  \begin{align*}
    \hat\ell^f((A',B');p)
    &= 2^{-|B|} \sum_{T\subseteq B} f(\hat A\triangle T\setminus B') + f(\hat A\triangle T\cup B')
    \\
    &= \frac 1 2 \; 2^{-|B|} \sum_{T\subseteq B} \biggl( f(\hat A\triangle T\setminus B') + f(\hat A\triangle T\cup B')
    \\[-12pt]
    & \hspace*{85pt}
      + f(\hat A\triangle (B\setminus T)\setminus B') + f(\hat A\triangle (B\setminus T)\cup B') \biggr)
      \numberthis\label{eq:tightness-double-counting}
    \\
    &\geq \frac 1 2 \; 2^{-|B|} \sum_{T\subseteq B} \biggl(
      f\Bigl((\hat A\triangle T\setminus B') \cup (\hat A\triangle (B\setminus T)\cup B')\Bigr)
    \\[-10pt]
    & \hspace*{85pt}
      + f\Bigl((\hat A\triangle T\setminus B') \cap (\hat A\triangle (B\setminus T)\cup B')\Bigr)
    \\[2pt]
    & \hspace*{85pt}
      + f\Bigl((\hat A\triangle (B\setminus T)\setminus B') \cup (\hat A\triangle T\cup B')\Bigr)
    \\[-2pt]
    & \hspace*{85pt}
      + f\Bigl((\hat A\triangle (B\setminus T)\setminus B') \cap (\hat A\triangle T\cup B')\Bigr)
      \biggr)
      \numberthis\label{eq:tightness-apply-submodularity}
    \\
    &= \frac 1 2 \; 2^{-|B|} \sum_{T\subseteq B} \biggl(
      f\bigl(\hat A\cup B\cup B'\bigr) + f\bigl(\hat A \setminus (B\cup B')\bigr)
    \\[-10pt]
    & \hspace*{85pt}
      f\bigl(\hat A\cup B\cup B'\bigr) + f\bigl(\hat A \setminus (B\cup B')\bigr)
      \biggr)      ~,
      \numberthis\label{eq:tightness-simplify}
    \\
    &= f\bigl(\hat A\cup B\cup B'\bigr) + f\bigl(\hat A \setminus (B\cup B')\bigr)
    \\
    &\geq f(B) + f(\emptyset)
      ~,
      \numberthis\label{eq:tightness-weak-increasing}
  \end{align*}
  where in eq.~\eqref{eq:tightness-apply-submodularity} we applied submodularity twice, to the first and fourth terms and to the second and third terms, and in eq.~\eqref{eq:tightness-weak-increasing} we used the fact that $f$ is increasing.
  We will show that for at least one $T\subseteq B$, the first and fourth terms in eq.~\eqref{eq:tightness-apply-submodularity} correspond to incomparable sets, meaning the inequality is actually tight.

  As $|B|\neq 1$, we must have $|B|\geq 2$.
  Let $i,j\in B$ be distinct elements such that $i\notin B'$.
  We therefore have two cases for $i$, and three for $j$, and for each we claim there exists $T\subseteq B$ such that $\hat A\triangle T\setminus B'$ and $\hat A\triangle (B\setminus T)\cup B'$ are incomparable:
  \begin{center}
    \begin{tabular}{R|CC}
      & i\in A' & i\notin A'
      \\
      \hline
      j\in A' & T=\{j\} & T=\{i,j\}
      \\
      j\in B' & T=B\setminus\{i\} & T=\{i\}
      \\
      j\notin A'\cup B' & T=\emptyset & T=\{i\}
    \end{tabular}
  \end{center}
  In particular, these choices give $i\in\hat A\triangle T$ and $i\notin\hat A\triangle (B\setminus T)$:
  in the first column, $i\in A'\setminus A \implies i\in \hat A$, so we ensure $i\notin T$, whereas we ensure $i\in T$ in the second column.
  Similarly, these choices give $j\notin\hat A\triangle T\setminus B'$ and $j\in\hat A\triangle (B\setminus T)\cup B'$:
  in the first row, $j\in \hat A$, so we ensure $j\in T$;
  in the second row, $j\in B'$, so we automatically have the condition;
  in the third row, $j\notin \hat A$, so we ensure $j\notin T$.
\end{proof}

\raf{Show also that $A,\{i\}$ is always weakly dominated by $A\cup\{i\},\emptyset$ or $A,\emptyset$.}

\raf{Other thoughts if we split this off into its own paper:}

\begin{enumerate}
\item 
  If modular on a sub lattice, what does that mean about abstentions wrt that sub lattice?  Maybe it’s not so bad to characterize the redundant reports as a function of f...
\item
  Related: what are examples of f functions in the wild?  Are they strict?
\item
  The link.  Would be super interesting to understand how to link properly when some reports are redundant.  Already not totally obvious for strictly modular and k=3... visualize!
\item
  Data set where Lh performs poorly wrt original target?
\end{enumerate}
